The Shape Matching Element Method: Direct Animation of Curved Surface Models

### 2021-04-27 paper/main.tex
add authors to main tex
Dave Changes
fixing typos

\documentclass[acmtog, anonymous,review, nonacm, balance = false]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

@@

>>>>

\documentclass[acmtog, anonymous=false, review=false, balance=false]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

\setcopyright{acmlicensed}\acmJournal{TOG}
\acmYear{2021}\acmVolume{40}\acmNumber{4}\acmArticle{69}\acmMonth{8} \acmDOI{10.1145/3450626.3459772}

@@

\author{Ty Trusty}

\email{trusty@cs.toronto.edu}
%
\author{Honglin Chen}

\email{chl9797@cs.toronto.edu}
%
%
\author{David I.W. Levin}
\email{diwlevin@cs.toronto.edu}

\affiliation{%
 \institution{University of Toronto}
 \streetaddress{40 St. George Street}
 \city{Toronto}
 \state{ON}
 \country{Canada}
 \postcode{M5S 2E4}
}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10010147</concept_id>
       <concept_desc>Computing methodologies</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10010147.10010371.10010352.10010379</concept_id>
       <concept_desc>Computing methodologies~Physical simulation</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies}
\ccsdesc[500]{Computing methodologies~Physical simulation}

%
% End generated code
%

\keywords{Physically-based animation, NURBS simulation, Meshless simulation, Lagrangian Mechanics}

### 2021-04-27 paper/sections/conclusion.tex
fixing typos

SEM itself has much room for improvement. 
First, while we focus on NURBS surfaces here, the only part of SEM that is NURBS specific is the shape matching operation.
We believe there is potential to allow mixed models (models which include polygonal meshes, particles, subdivision surfaces and NURBS)
by extending the range of shape matching operations used by the algorithm. The shape matching operation itself could be improved to be material-aware 
(to better handle heterogenous materials) or to be robust to noisy data (allowing direct simulation of scanned data).
\revise{In future work we also intend to explore using robust solvers or orthogonal bases for the shape matching step itself.} \ty{more can be said here I suppose. I tried pinv() on the beam, which surprisingly worked, but it's just unpredictable since it should still fail on my very basic 2D box example. Haven't done much experimentation beyond this though, so I just write future work}
It is also interesting to consider the relationship between our error term and the orthogonality constraint proposed by~\citet{Zhang:CompDynamics:2020}. 
Their approach could be used to remove this parameter from SEM entirely.
Finally, in this paper we have focused on applying SEM to physics-based animation, and there is a significant amount of additional work needed to 
extend SEM reliably into engineering applications.

>>>>

SEM itself has much room for improvement. 
First, while we focus on NURBS surfaces here, the only part of SEM that is NURBS specific is the shape matching operation.
We believe there is potential to allow mixed models (models which include polygonal meshes, particles, subdivision surfaces and NURBS)
by extending the range of shape matching operations used by the algorithm. The shape matching operation itself could be improved to be material-aware 
(to better handle heterogeneous materials) or to be robust to noisy data (allowing direct simulation of scanned data).
\revise{In future work we also intend to explore using robust solvers or orthogonal bases for the shape matching step itself.} \ty{more can be said here I suppose. I tried pinv() on the beam, which surprisingly worked, but it's just unpredictable since it should still fail on my very basic 2D box example. Haven't done much experimentation beyond this though, so I just write future work}
It is also interesting to consider the relationship between our error term and the orthogonality constraint proposed by~\citet{Zhang:CompDynamics:2020}. 
Their approach could be used to remove this term from SEM entirely.
Finally, in this paper we have focused on applying SEM to physics-based animation, and there is a significant amount of additional work needed to 
extend SEM reliably into engineering applications.

### 2021-04-27 paper/sections/intro.tex
fixing typos

NURBS (Non-uniform Rational \revise{B-Splines}) are a popular higher-order modelling primitive which 
are used for computer-aided design (CAD), computational fabrication and computer animation. 
NURBS primitives were the first geometric representation used for physics-based animation~\cite{10.1145/37402.37427}, yet,
despite over three decades of research, animation of curved models such as those built with NURBS, remains a challenge.

@@

We present the first truly meshless (no volumetric discretization required) algorithm for direct, nonlinear elastodynamic simulation of NURBs models. 
Our nonlinear elastodynamic simulation scheme requires only a boundary description of the object (we do not strictly require a solid model, see \reffig{teaser}) and
approriate physical parameters.  
Because we explicitly use the NURBS boundary representation in the simulation, it is straightforward to handle
Dirichlet and Neumann boundary conditions and to apply contact resolution. 
Crucially, because we broadly target animation and not necessarily simulation for engineering or manufacturing, we don't require that models
satisfy the rigourous geometric requirements common for these applications.

>>>>

NURBS (Non-uniform Rational \revise{B-Splines}) are a popular higher-order modelling primitive which 
are used for computer-aided design (CAD), computational fabrication and computer animation. 
NURBS primitives were the first geometric representation used for physics-based animation~\cite{10.1145/37402.37427}, yet,
despite over three decades of research, animation of curved models, such as those built with NURBS, remains a challenge.

@@

We present the first truly meshless (no volumetric discretization required) algorithm for direct, nonlinear elastodynamic simulation of NURBs models. 
Our nonlinear elastodynamic simulation scheme requires only a boundary description of the object (we do not strictly require a solid model, see \reffig{teaser}) and
appropriate physical parameters.  
Because we explicitly use the NURBS boundary representation in the simulation, it is straightforward to handle
Dirichlet and Neumann boundary conditions and to apply contact resolution. 
Crucially, because we broadly target animation and not necessarily simulation for engineering or manufacturing, we don't require that models
satisfy the rigorous geometric requirements common for these applications.

### 2021-04-27 paper/sections/methods.tex
fixing typos

%A pointform description of the general idea
%\begin{itemize}
%	\item A D-NURBS approach to simulation. Use a Jacobian to map from UV positions in a NURBS parametric space to world space positions on the surface. 
%	\item \textbf{Shape Matching}: From the sample world space positions on a single NURBS surface, we solve a least squares problem to fit a \textit{projection operator} to the surface. This projection operator maps monomials in the undeformed space, to the least squares estimate of the deformed positions for the given monomials.
%	\item With DNURBS and Shape Matching VEM we have two mappings: a UV to undeformed world space with our NURBS Jacobian, then an undeformed to deformed mapping with the projection operators. This lets us form a set of generalized coordinates in terms of our control points of the NURBS surface (section 1.4).
%	\item For an arbitrary position in space (not necessarily on the surface), we can build a projection operator unique to this point via a weighted sum of the projection operators of the projection operators for each NURBS (eq 5). From this we can reconstruct an estimate of the deformed position from the monomial basis defined using its undeformed position.
%	\item For any point in space, we need to compute a set of weights to the projection operators defined on the surfaces with the following criteria: the weights sum to 1, are nonzero, and depend on the distances to the surfaces (See section 1.8).
%	\item With this mapping for undeformed to deformed positions, this yields a simple definition for the deformation gradient (eq 12).
%	\item Armed with our deformation gradient, we next define the Lagrangian, but this basic form isn't sufficient. We augment our kinetic and potential energies with a stability term (ref VEM). This accounts for cases in which the position of the nodal values don't match their estimated positions defined by the projection operator mapping.
%	\item \textbf{Raycasting Quadrature}: To compute the volume integrals we make use of raycasting quadrature approach (ref) where we uniform sample a YZ grid and shoot rays in the X direction. We then use the points of intersection to define integrations bounds for 1D integrals along these rays, which we integrate numerically.
%	\item With our augmented lagrangian, we use the Euler-Lagrange equation to arrive at equations of motion. Our generalized coordinates give us direct updates on the control points of the NURBS. Currently we are using linearly-implicit Euler for time integration.
%	\item \textbf{Handling trimmed NURBS}: For this case we update the original control points, but we must only sample points within the boundary. To compute these points we use an approach similar to the raycasting quadrature but in 2D. In the paremtric domain for a T-NURBS we sample uniforly in the V direction and shoot rays along the U axis. At each region of intersection, we sample more values along the ray. The final result is a set of UV coordinates that respect the boundary defined by NURBS curves.
%\end{itemize}
\section{Overview}
The input to the Shape Matching Element Method (SEM)  is a NURBS \emph{boundary representation} of a volumetric object, along with a set of physical parameters 
(density, constitutive model, model parameters, etc). 
The boundary representation is composed of one or more (not necessarily explicitly connected) NURBS surface primitives that we will call \emph{parts}.
The output is an elastodynamic simulation. 
SEM consists of preprocessing and runtime simulation phases (~\refalg{sem_preprocess}, ~\refalg{sem_simulation}) and at no point do we need to generate a volumetric mesh of the input. 
In the \emph{preprocessing} stage we use raycasting to find quadrature points and weights for volumetric integration, as well as to compute part blending parameters. 
Finally we construct local shape matching operators and the mass matrix for our problem.

@@

Our kinematic map~(\refeq{kinematics}) requires blending weights in order to homogenize our per-part representation. 
Specifying weights manually would be antithetical to our goal of providing a seamless translation from surface representation to physics-based animation.
Automatic  weight computation is well-studied (see ~\citet{10.1145/2766952}, ~\citet{BBW:2011} and ~\citet{10.1145/2010324.1964968}), but typically requires solving 
an optimization problem that itself relies on a volumetric discretization. 
This can be avoided by using stochastic methods~\cite{Sawhney:2020:MCG} but these do not yet support the boundary conditions we require for our problem.

@@

\begin{equation}
\begin{aligned}
\mathbf{w}(\mathbf{X}_j) =  \arg\min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}_j) \mathbf{w}    \\
\textrm{s.t.} \quad & 0 \leq w_i \leq 1 \mbox{ }\forall i                    \\
                  &   \sum_i^n w_i = 1                      \\
\end{aligned}
\end{equation}

where $\Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_n}\right)$.

@@

\subsection{Choosing Deformation Origins}
\label{sec:origins}
% \dave{Ty: try to follow this formula when writing 
% \dave{maybe this should come after weight computation ?}
% (1) what is the purpose of this part of the method ?
% (2) what are the criteria for success
% (3) what is our assumed input 
% (4) what is the desired output
% (5) what's the algorith,
% (6) add proofs/diagrams/figures that show we satisfy (2). }
Each part, $\prt_i$, must be associated with a single deformation origin, $\bar{\refX}_j$, and each origin is computed as the centroid of the set of parts with which it is associated. 
While we constrain a part to be paired with only one deformation origin, an origin may be associated any number of parts.

@@

\begin{figure}[h]
    \includegraphics[width=\columnwidth]{figures/origin_network}
    \caption{Example structure for the network flow selection of deformation origins with a shape consisting of three parts. Red arrows indicate each parts origin association. \dave{make it clear edges of triangle are sep. parts, add capacities and costs}}
    \label{fig:network_flow} 
\end{figure}

@@

Given a quadrature point group, and an edge from the associated deformation origin to $\prt_i$, we set the edge cost to be the sum, 
over all quadrature points in the group of the blending weights associated with $\prt_i$.
This cost penalizes heavily ``connected'' origin candidates, encouraging sparsity during origin selection. 
This amounts to minimizing the following integer linear program:
\begin{equation}
\begin{aligned}
\mathbf{f} =  \arg\min_{\mathbf{f}} \quad & \mathbf{c}^T \mathbf{f}    \\
\textrm{s.t.} \quad & A_{\text{cons}} \mathbf{f} = 0  \\
                  & f_i \in \integer \mbox{ }, \forall i	\\
                  & 0 \leq f_i \leq u_i
\end{aligned}
\end{equation}
where $\mathbf{f}$ and $\mathbf{c}$ are the edge flow and cost values, respectively, assembled into vectors, and $A_{\text{cons}}$ forms the set of constraints satisfying the flow conservation constraint. 
We solve this using the dual simplex solver in MATLAB.
%\ty{we use the default in intlinprog, which should be dual-simplex} We choose $\XBs$ to be all the deformation origin candidates in $A$ that have an outgoing flow of at least $1$.

@@

We make a minor modification to this procedure to handle common geometric pathologies encountered in NURBS modelling, such as overlapping surfaces, by performing 
CSG unions while raycasting~\cite{ROTH1982109}. 
The output of this raycasting procedure is a set of quadrature points $\st{X}$ and weights $\st{v}$ that we use for integration and also as input to our 
deformation origin extraction (~\refsec{origins}) and blending weights computation (~\refsec{weights})
\reffig{raycasting} shows the convergence of this quadrature approach when used to compute the volume of a NURBS model. 
In principle we could apply detail-aware sampling methods~\cite{DBLP:journals/tog/WangFP12} to improve the robustness of raycasting quadrature even further,
but we found it unnecessary for the examples in this paper.
\begin{figure}
    \includegraphics[width=\columnwidth]{figures/raycasting_quadrature}
    \caption{When integrating physical quantities over the volumetric domain of an object, our raycasting quadrature is robust and converges to the groundtruth result as we increase the number of rays in use. }
    % \caption{Figure showing how raycasting quadrature works (top), convergence plot for volume integration (bottom}
    \label{fig:raycasting}
\end{figure}

>>>>

%A pointform description of the general idea
%\begin{itemize}
%	\item A D-NURBS approach to simulation. Use a Jacobian to map from UV positions in a NURBS parametric space to world space positions on the surface. 
%	\item \textbf{Shape Matching}: From the sample world space positions on a single NURBS surface, we solve a least squares problem to fit a \textit{projection operator} to the surface. This projection operator maps monomials in the undeformed space, to the least squares estimate of the deformed positions for the given monomials.
%	\item With DNURBS and Shape Matching VEM we have two mappings: a UV to undeformed world space with our NURBS Jacobian, then an undeformed to deformed mapping with the projection operators. This lets us form a set of generalized coordinates in terms of our control points of the NURBS surface (section 1.4).
%	\item For an arbitrary position in space (not necessarily on the surface), we can build a projection operator unique to this point via a weighted sum of the projection operators of the projection operators for each NURBS (eq 5). From this we can reconstruct an estimate of the deformed position from the monomial basis defined using its undeformed position.
%	\item For any point in space, we need to compute a set of weights to the projection operators defined on the surfaces with the following criteria: the weights sum to 1, are nonzero, and depend on the distances to the surfaces (See section 1.8).
%	\item With this mapping for undeformed to deformed positions, this yields a simple definition for the deformation gradient (eq 12).
%	\item Armed with our deformation gradient, we next define the Lagrangian, but this basic form isn't sufficient. We augment our kinetic and potential energies with a stability term (ref VEM). This accounts for cases in which the position of the nodal values don't match their estimated positions defined by the projection operator mapping.
%	\item \textbf{Raycasting Quadrature}: To compute the volume integrals we make use of raycasting quadrature approach (ref) where we uniform sample a YZ grid and shoot rays in the X direction. We then use the points of intersection to define integrations bounds for 1D integrals along these rays, which we integrate numerically.
%	\item With our augmented lagrangian, we use the Euler-Lagrange equation to arrive at equations of motion. Our generalized coordinates give us direct updates on the control points of the NURBS. Currently we are using linearly-implicit Euler for time integration.
%	\item \textbf{Handling trimmed NURBS}: For this case we update the original control points, but we must only sample points within the boundary. To compute these points we use an approach similar to the raycasting quadrature but in 2D. In the paremtric domain for a T-NURBS we sample uniforly in the V direction and shoot rays along the U axis. At each region of intersection, we sample more values along the ray. The final result is a set of UV coordinates that respect the boundary defined by NURBS curves.
%\end{itemize}
\section{Overview}
The input to the Shape Matching Element Method (SEM)  is a NURBS \emph{boundary representation} of a volumetric object, along with a set of physical parameters 
(density, constitutive model, model parameters, etc.). 
The boundary representation is composed of one or more (not necessarily explicitly connected) NURBS surface primitives that we will call \emph{parts}.
The output is an elastodynamic simulation. 
SEM consists of preprocessing and runtime simulation phases (\refalg{sem_preprocess}, ~\refalg{sem_simulation}) and at no point do we need to generate a volumetric mesh of the input. 
In the \emph{preprocessing} stage we use raycasting to find quadrature points and weights for volumetric integration, as well as to compute part blending parameters. 
Finally we construct local shape matching operators and the mass matrix for our problem.

@@

Our kinematic map~(\refeq{kinematics}) requires blending weights in order to homogenize our per-part representation. 
Specifying weights manually would be antithetical to our goal of providing a seamless translation from surface representation to physics-based animation.
Automatic  weight computation is well-studied (see ~\cite{10.1145/2766952}, ~\cite{BBW:2011} and ~\cite{10.1145/2010324.1964968}), but typically requires solving 
an optimization problem that itself relies on a volumetric discretization. 
This can be avoided by using stochastic methods~\cite{Sawhney:2020:MCG} but these do not yet support the boundary conditions we require for our problem.

@@

\begin{equation}
\begin{aligned}
\mathbf{w}(\mathbf{X}_j) =  \arg\min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}_j) \mathbf{w}    \\
\textrm{s.t.} \quad & 0 \leq w_i \leq 1 \mbox{ }\forall i                    \\
                  &   \sum_i^n w_i = 1                      \\
\end{aligned}
\end{equation}
where $\Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_n}\right)$.

@@

\subsection{Choosing Deformation Origins}
\label{sec:origins}
% \dave{Ty: try to follow this formula when writing 
% \dave{maybe this should come after weight computation ?}
% (1) what is the purpose of this part of the method ?
% (2) what are the criteria for success
% (3) what is our assumed input 
% (4) what is the desired output
% (5) what's the algorith,
% (6) add proofs/diagrams/figures that show we satisfy (2). }
Each part, $\prt_i$, must be associated with a single deformation origin, $\bar{\refX}_j$, and each origin is computed as the centroid of the set of parts with which it is associated. 
While we constrain a part to be paired with only one deformation origin, an origin may be associated with any number of parts.

@@

\begin{figure}[h]
    \includegraphics[width=\columnwidth]{figures/origin_network}
    \caption{Example structure for the network flow selection of deformation origins with a shape consisting of three parts. Red arrows indicate each part's origin association. \dave{make it clear edges of triangle are sep. parts, add capacities and costs}}
    \label{fig:network_flow} 
\end{figure}

@@

Given a quadrature point group, and an edge from the associated deformation origin to $\prt_i$, we set the edge cost to be the sum, 
over all quadrature points in the group of the blending weights associated with $\prt_i$.
This cost penalizes heavily ``connected'' origin candidates, encouraging sparsity during origin selection. 
This amounts to minimizing the following integer linear program:
\begin{equation}
\begin{aligned}
\mathbf{f} =  \argmin_{\mathbf{f}} \quad & \mathbf{c}^T \mathbf{f}    \\
\textrm{s.t.} \quad & A_{\text{cons}} \mathbf{f} = 0  \\
                  & f_i \in \integer \mbox{ }, \forall i	\\
                  & 0 \leq f_i \leq u_i
\end{aligned}
\end{equation}
where $\mathbf{f}$ and $\mathbf{c}$ are the edge flow and cost values, respectively, assembled into vectors, and $A_{\text{cons}}$ forms the set of constraints satisfying the flow conservation constraint. 
We solve this using the dual simplex solver in MATLAB.
%\ty{we use the default in intlinprog, which should be dual-simplex} We choose $\XBs$ to be all the deformation origin candidates in $A$ that have an outgoing flow of at least $1$.

@@

We make a minor modification to this procedure to handle common geometric pathologies encountered in NURBS modelling, such as overlapping surfaces, by performing 
CSG unions while raycasting~\cite{ROTH1982109}. 
The output of this raycasting procedure is a set of quadrature points $\st{X}$ and weights $\st{v}$ that we use for integration and also as input to our 
deformation origin extraction (\refsec{origins}) and blending weights computation (\refsec{weights})
\reffig{raycasting} shows the convergence of this quadrature approach when used to compute the volume of a NURBS model. 
In principle we could apply detail-aware sampling methods~\cite{DBLP:journals/tog/WangFP12} to improve the robustness of raycasting quadrature even further,
but we found it unnecessary for the examples in this paper.
\begin{figure}
    \includegraphics[width=\columnwidth]{figures/raycasting_quadrature}
    \caption{When integrating physical quantities over the volumetric domain of an object, our raycasting quadrature is robust and converges to the groundtruth result as we increase the number of rays in use. }
    % \caption{Figure showing how raycasting quadrature works (top), convergence plot for volume integration (bottom}
    \label{fig:raycasting}
\end{figure}

### 2021-04-27 paper/sections/results.tex
fixing typos

  \caption{Performance of, and parameters for the Shape Matching Element Method on all examples. All wall-clock timings are reported in seconds, physical parameters are

@@

\revise{Additionally, we perform a twisting beam experiment on a 22 part beam model. Despite the relatively small number of NURBS patches, we see in \reffig{beam_twist} that the quadratic polynonmials permit complex deformation.}

@@

SEM also supports the simulation of objects made up of heterogenous materials. By specifying different material properties inside nested 
parts we are able to simulate a drag racing wheel with a metal hub and soft rubber tread~(\reffig{tire}).

@@

%editing example
For downstream applications, the output of  SEM simulation is itself editable in a NURBS modelling program (see \reffig{edit}). 
This can be useful for visual effects pipelines, allowing artists to easily post process animations using the same tools in which they were created.
\begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/export_iges.pdf}
  \caption{Here we show the output of an SEM animation loaded into Rhinoceros 3D 7 for additional editing.\ty{note: initial version of figure. Currently remaking it :)}}
  \label{fig:edit}
\end{figure}

@@

In \reffig{teaser} we directly simulate the NURBS surface model of a bouncy castle under a periodic wind force. 
In \reffig{starship} we show an animation of the Space-V starship landing on Mars in a graceful way. 
This example shows the benefit of NURBS modeling, a relatively small number of primitives can represent a complex shape.
SEM successfully handles the stiff materials, complicated geometry and non-trivial collisions in this scene.

>>>>

  \caption{Performance and parameters for the Shape Matching Element Method on all examples. All wall-clock timings are reported in seconds, physical parameters are

@@

\revise{Additionally, we perform a twisting beam experiment on a 22 part beam model. Despite the relatively small number of NURBS patches, we see in \reffig{beam_twist} that the quadratic polynomials permit complex deformation.}

@@

SEM also supports the simulation of objects made up of heterogeneous materials. By specifying different material properties inside nested 
parts we are able to simulate a drag racing wheel with a metal hub and soft rubber tread~(\reffig{tire}).

@@

%editing example
For downstream applications, the output of  SEM simulation is itself editable in a NURBS modelling program (see \reffig{edit}). 
This can be useful for visual effects pipelines, allowing artists to easily post-process animations using the same tools in which they were created.
\begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/export_iges.pdf}
  \caption{Here we show the output of an SEM animation loaded into Rhinoceros 3D 7 for additional editing.\ty{note: initial version of figure. Currently remaking it :)}}
  \label{fig:edit}
\end{figure}

@@

In \reffig{teaser} we directly simulate the NURBS surface model of a bouncy castle under a periodic wind force. 
In \reffig{starship} we show an animation of the Space-V starship landing on Mars in a graceful way. 
This example shows the benefit of NURBS modeling: a relatively small number of primitives can represent a complex shape.
SEM successfully handles the stiff materials, complicated geometry and non-trivial collisions in this scene.
